Reproductive Allocation Calculations

### 2015-03-24 leaflifespan.Rmd
working to streamline script, use functions for plotting, models

LMA <- read.csv("data/LMA_short.csv",header=TRUE,as.is=TRUE,stringsAsFactors=FALSE)
#InvestmentSummary <- remake::make("ReproductionAllocation_all")
#names(InvestmentSummary)[1] <- "tag_ID"

@@

# where counts did not previously exists, set as 0
for(v in c("lvs_start_count","lvs_end_count","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leaf_lifespan[[v]])
  leaf_lifespan[[v]][i] <- 0
}

@@

# if possible use data for that individual, if it exists
i <- !is.na(raw_leaf_lifespan$mm_lvs)
raw_leaf_lifespan$count_per_length[i] <- (raw_leaf_lifespan$mm_lvs/raw_leaf_lifespan$count_lvs_spec)[i]

@@

"lvs_start_length""lvs_end_length" "lvs_new_length", ,"mm_lvs_spec","count_lvs_spec"
# #turning NAs into zeros
# colnames(leaf_lifespan)
# leaf_lifespan$lvs_start_length[is.na(leaf_lifespan$lvs_start_length)] <-0
# leaf_lifespan$lvs_start_count[is.na(leaf_lifespan$lvs_start_count)] <-0
# leaf_lifespan$lvs_end_length[is.na(leaf_lifespan$lvs_end_length)] <-0
# leaf_lifespan$lvs_end_count[is.na(leaf_lifespan$lvs_end_count)] <-0
# leaf_lifespan$lvs_new_length[is.na(leaf_lifespan$lvs_new_length)] <-0
# leaf_lifespan$lvs_new_count[is.na(leaf_lifespan$lvs_new_count)] <-0
# leaf_lifespan$new_but_shed_count[is.na(leaf_lifespan$new_but_shed_count)] <-0
# leaf_lifespan$mm_lvs_spec[is.na(leaf_lifespan$mm_lvs_spec)] <-0
# leaf_lifespan$count_lvs_spec[is.na(leaf_lifespan$count_lvs_spec)] <-0

# specific leaves per length data
leaf_lifespan_specific <- subset(leaf_lifespan,mm_lvs_spec!="0")
leaf_lifespan_specific$count_per_length <- leaf_lifespan_specific$count_lvs_spec/leaf_lifespan_specific$mm
leaf_lifespan_specific$leaf_start_total <- leaf_lifespan_specific$lvs_start_count + (leaf_lifespan_specific$count_per_length*leaf_lifespan_specific$lvs_start_length)
leaf_lifespan_specific$leaf_end_total <- leaf_lifespan_specific$lvs_end_count + (leaf_lifespan_specific$count_per_length*leaf_lifespan_specific$lvs_end_length)
leaf_lifespan_specific$leaf_new_total <- leaf_lifespan_specific$lvs_new_count + leaf_lifespan_specific$new_but_shed_count + (leaf_lifespan_specific$count_per_length*leaf_lifespan_specific$lvs_new_length)
head(leaf_lifespan_specific)

#generic leaves per length data
leaves_per_length$count_per_length <- leaves_per_length$leaf_count/leaves_per_length$length_mm
per_length_spp <- aggregate(count_per_length ~ species, leaves_per_length, FUN=mean)

leaf_lifespan_generic <- subset(leaf_lifespan,mm_lvs_spec=="0")
leaf_lifespan_generic <- merge(leaf_lifespan_generic, per_length_spp, by.x="species", by.y="species")
leaf_lifespan_generic$leaf_start_total <- leaf_lifespan_generic$lvs_start_count + (leaf_lifespan_generic$count_per_length*leaf_lifespan_generic$lvs_start_length)
leaf_lifespan_generic$leaf_end_total <- leaf_lifespan_generic$lvs_end_count + (leaf_lifespan_generic$count_per_length*leaf_lifespan_generic$lvs_end_length)
leaf_lifespan_generic$leaf_new_total <- leaf_lifespan_generic$lvs_new_count + leaf_lifespan_generic$new_but_shed_count + (leaf_lifespan_generic$count_per_length*leaf_lifespan_generic$lvs_new_length)
leaf_lifespan_generic <- subset(leaf_lifespan_generic,leaf_new_total>"0")
head(leaf_lifespan_generic)

#combine two files into 1
leaf_counts <- bind_rows(leaf_lifespan_specific, leaf_lifespan_generic)
head(leaf_counts)

#calculating leaf lifespan
leaf_counts$LL_bd <- (leaf_counts$leaf_start_total)/(((leaf_counts$leaf_start_total)-(leaf_counts$leaf_end_total))+(leaf_counts$leaf_new_total)/2)
leaf_counts$LL_death <- (leaf_counts$leaf_start_total)/((leaf_counts$leaf_start_total)-(leaf_counts$leaf_end_total))
leaf_counts$LL_birth <- (leaf_counts$leaf_start_total)/(leaf_counts$leaf_new_total)

leaf_counts$LL_death[which(is.infinite(leaf_counts$LL_death))] <- NA
leaf_counts$LL_bd[which(is.infinite(leaf_counts$LL_death))] <- NA
leaf_counts$LL_birth[which(is.infinite(leaf_counts$LL_death))] <- NA

write.csv(leaf_counts, file = "leaf_counts.csv")

#adding leaf, stem weight data from InvestmentSummary to leaf_counts dataframe
leaf_counts <- left_join(leaf_counts, InvestmentSummary, by = "tag_ID")

#adding columns to leaf_counts to look at change across year
leaf_counts$change_shoot_diam <- leaf_counts$d_end - leaf_counts$d_start
leaf_counts$change_basal_diam <- leaf_counts$FinalBasalDiam - leaf_counts$StartBasalDiam
leaf_counts$change_shoot_area <- (3.14*((leaf_counts$d_end/2)^2)) - (3.14*((leaf_counts$d_start/2)^2))
leaf_counts$change_basal_area <- (3.14*((leaf_counts$FinalBasalDiam/2)^2)) - (3.14*((leaf_counts$StartBasalDiam/2)^2))

#removng crazy numbers
leaf_counts <- subset(leaf_counts,LL_bd>0)
leaf_counts <- subset(leaf_counts,LL_bd<6)
leaf_counts <- subset(leaf_counts,LL_birth<8)
leaf_counts <- subset(leaf_counts,LL_death<8)

#leaf lifespan summary by species, age
LLSummary <- leaf_counts %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length)

@@

leaf_counts_nobabies <- subset(leaf_counts, age>2)
LMA_nobabies <- subset(LMA, age>2)

#leaf lifespan summary by species
LL_spp <- leaf_counts_nobabies %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

#LMA summary by species
LMA_spp <- LMA_nobabies %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LMA)
names(LMA_spp)<-c("species", "LMA_mean","LMA_se","LMA_length")

@@

#plotting and stats
#LL vs LMA
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","light blue")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia","Boronia","Conospermum","Epacris", "Grevillea buxifolia","Grevillea speciosa","Hakea","Hemigenia","Leucopogon","Persoonia", "Petrophile", "Phyllota", "Pimelea", "Pultanaea")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU") 
labels.age <- c("1.4","2.4","5","7","9","32")

#remove HATE, babies
LL_nobabies <- subset(LLSummary, age > 2)
LL_noHate <- subset(LLSummary, species != "HATE")
LL_noHatebabies <- subset(LL_nobabies, species != "HATE")
```

@@

For 2.4 year old plants

```{r, echo=FALSE}
plot(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==2.4), xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1)
mod <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==2.4))
summary(mod)
abline(mod)
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_basal_diam ~ leaf_new_total, leaf_counts, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in basal diameter")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_basal_diam+1) + species +age, leaf_counts)
summary(mod)

plot(change_shoot_diam ~ leaf_new_total, leaf_counts, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in diameter of leader")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_shoot_diam+10) + species +age, data= subset(leaf_counts, !is.na(change_shoot_diam)))
summary(mod)
```

@@

col.LMA <-c("black","purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("all except HATE","age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "including HATE")
plot(LL_birth_mean ~ LMA_mean, LLSummary, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, LL_noHate, pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(LLSummary, species=="HATE"), pch=17)
mod <- lm(LL_birth_mean ~ LMA_mean, LLSummary)
abline(mod, col="red", lty=2)
mod_no_HATE <- lm(LL_birth_mean ~ LMA_mean, LL_noHate)
abline(mod_no_HATE, lwd=2)
mod_no_HATE5 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==5))
abline(mod_no_HATE5, col="light green", lwd=1.5)
mod_no_HATE7 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==7))
abline(mod_no_HATE7, col="orange", lwd=1.5)
mod_no_HATE9 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==9))
abline(mod_no_HATE9, col="blue", lwd=1.5)
mod_no_HATE2.4 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==2.4))
abline(mod_no_HATE2.4, col="pink", lwd=1.5)
mod_no_HATE1 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age<2.4))
abline(mod_no_HATE1, col="purple", lwd=1.5)
mod_no_HATE32 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LL_noHate,age==32))
abline(mod_no_HATE32, col="light blue", lwd=1.5)

@@

```{r, echo=FALSE,results='hide', message=FALSE}
mod <- lm(LL_birth ~ age +species + age*species, leaf_counts_nobabies)
summary(mod)

leaf_counts %>% group_by(species) %>%
    summarise(mean(age))

plot_LLvAge <- function(spp) {
  plot(LL_birth ~ age, data=subset(leaf_counts,species==spp), xlab="age", ylab="leaf lifespan (leaf births)", cex.axis=.8, las=1, 
     main=labels.spp[spp], xlim=c(1,35), ylim=c(0,8), log="x")  
}

for(n in names(labels.spp)) {
  plot_LLvAge(n)
}
```

@@

```{r, echo=FALSE, message=FALSE}
plot(LMA_mean ~ as.factor(species), LL_nobabies, xlab="species", ylab="LMA", cex.axis=.8, las=1)
mod <- lm(LMA ~ age +species + age*species, LMA)
summary(mod)

plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="BAER"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Banksia ericifolia")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="BOLE"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Boronia ledifolia")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="COER"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Conospermum ericifolium")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="EPMI"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Epacris microphylla")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="GRBU"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Grevillea buxifolia")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="GRSP"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Grevillea speciosa")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="HATE"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Hakea teretifolia")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="HEPU"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Hemigenia purpurea")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="LEES"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Leucopogon esquamatus")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="PELA"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Persoonia lanceolata")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="PEPU"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Petrophile pulchella")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="PHPH"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Phyllota phylicoides")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="PILI"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Pimelea linifolia")
plot(LMA ~ as.factor(age), LMA, data=subset(LMA,species=="PUTU"), xlab="age", ylab="LMA", cex.axis=.8, las=1, main="Pultenaea tuberculata")
```

#models with leaf counts per age
The number of leaves on a leader is slightly affected by age however, with a decrease in leaf count along the leader with increasing age. This is what gives the perception of fewer leaves in older plants.

QUESTION: Why does age need to be a factor for the plots to work?
```{r, echo=FALSE,results='hide', message=FALSE}

plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="BAER"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Banksia ericifolia")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="BOLE"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Boronia ledifolia")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="COER"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Conospermum ericifolium")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="EPMI"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Epacris microphylla")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="GRBU"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Grevillea buxifolia")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="GRSP"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Grevillea speciosa")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="HATE"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Hakea teretifolia")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="HEPU"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Hemigenia purpurea")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="LEES"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Leucopogon esquamatus")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="PELA"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Persoonia lanceolata")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="PEPU"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Petrophile pulchella")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="PHPH"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Phyllota phylicoides")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="PILI"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Pimelea linifolia")
plot(leaf_end_total ~ as.factor(age), leaf_counts, data=subset(leaf_counts,species=="PUTU"), xlab="age", ylab="leaf counts", cex.axis=.8, las=1, main="Pultenaea tuberculata")

mod <- lm(leaf_end_total ~ age +species +age*species, leaf_counts_nobabies)
summary(mod)
```

#models with shoot extenstion by age

###combine shoot extension and investment files, by tag_ID

```{r, echo=FALSE,results='hide', message=FALSE}
plot(new_length_mean ~ age, LLSummary, col=col.spp[as.factor(species)])
legend("topright",col=col.spp, labels.spp, pch=16, cex=1.4)

>>>>

LMA <- read.csv("data/LMA_short.csv",header=TRUE,as.is=TRUE,stringsAsFactors=FALSE)%>%
  filter(species!="" & species!=" ")
#InvestmentSummary <- remake::make("ReproductionAllocation_all")
#names(InvestmentSummary)[1] <- "tag_ID"

@@

# where counts did not previously exists, set as 0; CHECK WITH DANIEL THAT "V" IS A SET "NAME" IN THESE FUNCTIONS
for(v in c("lvs_start_count","lvs_end_count","lvs_new_count","new_but_shed_count")) {
  i <- is.na(leaf_lifespan[[v]])
  leaf_lifespan[[v]][i] <- 0
}

@@

# if possible use data for that individual, if it exists; this code substitutes in species count per length #s for individuals where this data is available
i <- !is.na(raw_leaf_lifespan$mm_lvs)
raw_leaf_lifespan$count_per_length[i] <- (raw_leaf_lifespan$count_lvs_spec/raw_leaf_lifespan$mm_lvs)[i]

#this is the end of the section Daniel helped me rewrite - next section is mine

#get total leaf counts
for(v in c("lvs_start_length","lvs_start_count","lvs_end_length","lvs_end_count","lvs_new_length","lvs_new_count","new_but_shed_count")) {
  i <- is.na(raw_leaf_lifespan[[v]])
  raw_leaf_lifespan[[v]][i] <- 0
  }
raw_leaf_lifespan$lvs_start <- raw_leaf_lifespan$lvs_start_count + (raw_leaf_lifespan$lvs_start_length * raw_leaf_lifespan$count_per_length)
raw_leaf_lifespan$lvs_end <- raw_leaf_lifespan$lvs_end_count + (raw_leaf_lifespan$lvs_end_length * raw_leaf_lifespan$count_per_length)
raw_leaf_lifespan$lvs_new <- raw_leaf_lifespan$lvs_new_count + (raw_leaf_lifespan$lvs_new_length * raw_leaf_lifespan$count_per_length)

#calculating leaf lifespan
raw_leaf_lifespan$LL_bd <- (raw_leaf_lifespan$lvs_start)/(((raw_leaf_lifespan$lvs_start)-(raw_leaf_lifespan$lvs_end))+(raw_leaf_lifespan$lvs_new)/2)
raw_leaf_lifespan$LL_death <- (raw_leaf_lifespan$lvs_start)/((raw_leaf_lifespan$lvs_start)-(raw_leaf_lifespan$lvs_end))
raw_leaf_lifespan$LL_birth <- (raw_leaf_lifespan$lvs_start)/(raw_leaf_lifespan$lvs_new)

raw_leaf_lifespan$LL_death[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA
raw_leaf_lifespan$LL_bd[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA
raw_leaf_lifespan$LL_birth[which(is.infinite(raw_leaf_lifespan$LL_death))] <- NA

write.csv(raw_leaf_lifespan, file = "raw_leaf_lifespan.csv")
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
#adding leaf, stem weight data from InvestmentSummary to raw_leaf_lifespan dataframe; can't RUN LINES 95-108 UNTIL INVESTMENT SUMMARY OPEN - THEN PROBLEM WITH PLYR
#raw_leaf_lifespan <- left_join(raw_leaf_lifespan, InvestmentSummary, by = "tag_ID")

#adding columns to raw_leaf_lifespan to look at change across year
raw_leaf_lifespan$change_shoot_diam <- raw_leaf_lifespan$d_end - raw_leaf_lifespan$d_start
#raw_leaf_lifespan$change_basal_diam <- raw_leaf_lifespan$FinalBasalDiam - raw_leaf_lifespan$StartBasalDiam
raw_leaf_lifespan$change_shoot_area <- (3.14*((raw_leaf_lifespan$d_end/2)^2)) - (3.14*((raw_leaf_lifespan$d_start/2)^2))
#raw_leaf_lifespan$change_basal_area <- (3.14*((raw_leaf_lifespan$FinalBasalDiam/2)^2)) - (3.14*((raw_leaf_lifespan$StartBasalDiam/2)^2))

#removng crazy numbers; probably better to do this for each analysis than overall, because removing indiviiduals from all three at once
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_bd>0)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_bd<6)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_birth<8)
#raw_leaf_lifespan <- subset(raw_leaf_lifespan,LL_death<8)

#creating summary files
#leaf lifespan summary by species, age
LLSummary <- raw_leaf_lifespan %>%
  filter(LL_bd>0 & LL_bd<6 & LL_birth<8 & LL_death<8) %>%
  group_by(species, age) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth,new_length, lvs_end)

@@

#leaf lifespan summary by species

LL_spp <- raw_leaf_lifespan %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LL_bd, LL_death,LL_birth)

#LMA summary by species
LMA_spp <- LMA %>%
  filter(age>2) %>%
  group_by(species) %>%
  summarise_each(funs(mean, se, length), LMA)
names(LMA_spp)<-c("species", "LMA_mean","LMA_se","LMA_length")

@@

#plotting and stats
#creating lists of information for plotting symbols, colors, labels
pch.spp <- c(0,1,2,15,16,17,18,0,1,2,15,16,17,18)
col.spp <- c("red", "dark green", "blue", "purple", "orange", "light green","salmon","turquoise","grey", "black","light blue","brown","yellow","salmon")
col.age <-c("purple","pink","green","orange","blue","light blue")
pch.age <- c(1,1,16,16,16,16)
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU") 
labels.age <- c("1.4","2.4","5","7","9","32")

#LL vs LMA
#remove HATE, babies
#instead of separate dataframes, can probably incorporate this information directly into plotting and analysis commands
#LL_nobabies <- subset(raw_leaf_lifespan, age>2)
#LMA_nobabies <- subset(LMA, age>2)
#LL_noHate <- subset(LLSummary, species != "HATE")
#LL_noHatebabies <- subset(LL_nobabies, species != "HATE")
```

@@

For 7 year old plants
```{r, echo=FALSE}
plot(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7), xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA", las=1, pch=16)
text(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7), labels=species, pos=1, offset=-0.9, cex=0.9,col="blue")
text(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7 & species=="HATE"), labels=species, pos=1, offset=-0.9, cex=0.9,col="red")
mod <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,species!="HATE" & age==7))
summary(mod)
abline(mod)
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_basal_diam ~ leaf_new_total, raw_leaf_lifespan, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in basal diameter")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_basal_diam+1) + species +age, raw_leaf_lifespan)
summary(mod)

plot(change_shoot_diam ~ leaf_new_total, raw_leaf_lifespan, log="xy", col=col.age[as.factor(age)], pch=16, xlab="leaf number on leader", ylab="change in diameter of leader")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
mod <- lm(log(leaf_new_total) ~ log(change_shoot_diam+10) + species +age, data= subset(raw_leaf_lifespan, !is.na(change_shoot_diam)))
summary(mod)
```

@@

col.LMA <-c("black","purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("all except HATE","age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "including HATE")
plot(LL_birth_mean ~ LMA_mean, LLSummary, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(LLSummary, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(LLSummary, species=="HATE"), pch=17)
#need log transformation to get these to plot onto axes
mod <- lm(LL_birth_mean ~ LMA_mean, LLSummary)
abline(mod, col="red", lty=2)
mod_no_HATE <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary, species!="HATE"))
abline(mod_no_HATE, lwd=2)
mod_no_HATE5 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==5&species!="HATE"))
abline(mod_no_HATE5, col="light green", lwd=1.5)
mod_no_HATE7 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==7&species!="HATE"))
abline(mod_no_HATE7, col="orange", lwd=1.5)
mod_no_HATE9 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==9&species!="HATE"))
abline(mod_no_HATE9, col="blue", lwd=1.5)
mod_no_HATE2.4 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==2.4&species!="HATE"))
abline(mod_no_HATE2.4, col="pink", lwd=1.5)
mod_no_HATE1 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age<2.4&species!="HATE"))
abline(mod_no_HATE1, col="purple", lwd=1.5)
mod_no_HATE32 <- lm(LL_birth_mean ~ LMA_mean, data=subset(LLSummary,age==32&species!="HATE"))
abline(mod_no_HATE32, col="light blue", lwd=1.5)

@@

```{r, echo=FALSE,results='hide', message=FALSE}
mod <- lm(LL_birth ~ age + species +age*species, data=subset(raw_leaf_lifespan,age>3 & LL_birth!="Inf"))
summary(mod)

#WHY ARE THE FOLLOWING TWO LINES OF CODE NEEDED?
raw_leaf_lifespan %>% group_by(species) %>%
    summarise(mean(LL_birth))

#where does the [spp] comes from? This is the first place in the script it is used like this - I assume it could be any name?
plot_LLvAge <- function(spp) {
  plot(LL_birth ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf lifespan (leaf births)", cex.axis=.8, las=1, 
     main=labels.spp[spp], xlim=c(1,35), ylim=c(0,8), log="x", pch=16)  
  points(LL_birth_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((LL_birth_mean+LL_birth_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((LL_birth_mean-LL_birth_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

for(n in names(labels.spp)) {
  plot_LLvAge(n)
}

```

@@

```{r, echo=FALSE, message=FALSE}
#plot(LMA_mean ~ as.factor(species), data=subset(LLSummary,age>3), xlab="species", ylab="LMA", cex.axis=.8, las=1)
#mod <- lm(LMA ~ age +species + age*species, LMA)
#summary(mod)

LMA %>% group_by(species) %>%
  summarise(mean(LMA))

labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")

plot_LMAvAge <- function(spp) {
  plot(LMA ~ age, data=subset(LMA,species==spp), xlab="age", ylab="LMA", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(LMA_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((LMA_mean+LMA_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((LMA_mean-LMA_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
  #I want to add a line of code that plots over the means for each age. I did this above by calling in another file, but wondering about on-the-spot calculations
}

mod_LMAvAge <- function(spp) {
  lm(LMA ~ age, data=subset(LMA,species==spp))
}

for(n in names(labels.spp)) {
  plot_LMAvAge(n)
  mod_LMAvAge(n)
  summary(mod_LMAvAge(n))
}

```

#models with leaf counts per age
The number of leaves on a leader is slightly affected by age however, with a decrease in leaf count along the leader with increasing age. This is what gives the perception of fewer leaves in some older plants.

labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")

plot_lfcountvAge <- function(spp) {
  plot(lvs_end ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leaf count along shoot", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(lvs_end_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((lvs_end_mean+lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((lvs_end_mean-lvs_end_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

mod_lfcountvAge <- function(spp) {
  lm(lvs_end ~ age, data=subset(raw_leaf_lifespan,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_lfcountvAge(n)
  mod_lfcountvAge(n)
  summary(mod_lfcountvAge(n))
}

```

#models with shoot extenstion by age
These models show that for some species, shoot extension decreases enormously in the older plants. This effect is larger than with number of leaves, indicating that some of these individuals have more leaves per length as they get older

```{r, echo=FALSE,results='hide', message=FALSE}
labels.spp <- c("Banksia ericifolia","Boronia ledifolia","Conospermum ericifolium","Epacris micraphylla", "Grevillea buxifolia","Grevillea speciosa","Hakea teretifolia","Hemigenia purpurea","Leucopogon esquamatus","Persoonia lanceolata", "Petrophile puchella", "Phyllota phyllicoides", "Pimelea linifolia", "Pultanaea tuberculata")
names(labels.spp) <- c("BAER","BOLE","COER","EPMI","GRBU","GRSP","HATE","HEPU","LEES","PELA","PEPU","PHPH","PILI","PUTU")

plot_newlengthvAge <- function(spp) {
  plot(new_length ~ age, data=subset(raw_leaf_lifespan,species==spp), xlab="age", ylab="leader shoot extention", cex.axis=.8, las=1, pch=16, log="x",
  main=labels.spp[spp]) 
  points(new_length_mean ~ age, data=subset(LLSummary, species==spp), cex=1.3, pch=8)
  points((new_length_mean+new_length_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=25)
  points((new_length_mean-new_length_se) ~ age, data=subset(LLSummary, species==spp), cex=1.5, pch=24)
}

mod_newlengthvAge <- function(spp) {
  lm(new_length ~ age, data=subset(raw_leaf_lifespan,age>2.4 & species==spp))
}

for(n in names(labels.spp)) {
  plot_newlengthvAge(n)
  mod_newlengthvAge(n)
  summary(mod_newlengthvAge(n))
}
```

#plots regressing variables against one another
